### 2017-02-20 writing/paper/sections/coincidences.tex
Finish writing paper v1

Coincidences among small mesh patterns have previously been considered by 
\textcite{DBLP:journals/corr/ClaessonTU14}, in which the authors use the Simultaneous Shading Lemma,
a closure result and one worked out special case to fully classify coincidences among mesh patterns
of length 2.

@@

We will consider the avoidance sets \(\av{\pi,p}\) where \(\pi\) is a classical pattern of length 3
and \(p\) is a mesh pattern of length 2 in order to establish sufficient conditions for two such
sets to be coincident. We will fix \(\pi\) in order to define these coincidences and say that \(\pi\)
is the \emph{dominating pattern}.

In order to describe the rules it is useful to have a notion for inserting points, ascents, and descents
into a mesh pattern.
\begin{definition}
\label{def:ap}
Let \(p=(\tau,R)\) be a mesh pattern of length \(n\) such that \(\boks{i,j}\notin R\). We define
a mesh pattern \(p^\boks{i,j} = (\tau^\prime,R^\prime)\) of length \(n+1\) as the pattern where a 
point is \emph{inserted} into the box \(\boks{i,j}\) in \(G(p)\). Formally the new underlying classical
pattern is defined by 
\begin{equation*}
\tau^\prime(k) = \begin{cases}
    j+1 & \text{if } k = i+1\\
    \tau(k) & \text{if } \tau(k)\le j \text{ and }k\le i\\
    \tau(k)+1 & \text{if } \tau(k)> j \text{ and }k\le i\\
    \tau(k-1) & \text{if } \tau(k)\le j \text{ and }k> i+1\\
    \tau(k-1)+1 & \text{if } \tau(k)> j \text{ and }k> i+1\\
\end{cases} 
\end{equation*}
While the mesh becomes
\begin{equation*}
\begin{aligned}
R^\prime=&\{\boks{k,\ell}\mid k\le i, \ell\le j, \boks{k,\ell}\in R\} \cup\\
&\{\boks{k,\ell}\mid k\le i, \ell > j, \boks{k,\ell-1}\in R\}\cup\\
&\{\boks{k,\ell}\mid k > i, \ell \le j, \boks{k-1,\ell}\in R\}\cup\\
&\{\boks{k,\ell}\mid k > i, \ell > j, \boks{k-1,\ell-1}\in R\}\\
\end{aligned}
\end{equation*}
\end{definition}
In addition, we give the following definitions:
\begin{definition}
Let \(p=(\tau,R)\) be a mesh pattern of length \(n\) such that \(\boks{i,j}\notin R\) and 
\(p^\boks{i,j} = (\tau^\prime,R^\prime)\) to be as defined in \DefinitionRef{def:ap}. We define
the following five modifications of the mesh patterns of the same length as \(p^\boks{i,j}\). 
\begin{align*}
p^\boksup{i,j} &= (\tau^\prime, R^\prime \cup \{\boks{i,j+1},\boks{i+1,j+1}\})\\
p^\boksright{i,j} &= (\tau^\prime, R^\prime \cup \{\boks{i+1,j},\boks{i+1,j+1}\})\\
p^\boksdown{i,j} &= (\tau^\prime, R^\prime \cup \{\boks{i,j},\boks{i+1,j}\})\\
p^\boksleft{i,j} &= (\tau^\prime, R^\prime \cup \{\boks{i,j},\boks{i,j+1}\})\\
\end{align*}
\end{definition}
Informally, these are considering the topmost, rightmost, leftmost, or bottommost point in \(\boks{i,j}\).
We allow composition of these modifications, and collect the resulting mesh patterns in a set
\begin{equation*}
p^\boksall{i,j}=\left\{
p^\boks{i,j},
p^\boksright{i,j},
p^\boksrightup{i,j},
p^\boksrightdown{i,j},
p^\boksleft{i,j},
p^\boksleftup{i,j},
p^\boksleftdown{i,j},
p^\boksup{i,j},
p^\boksdown{i,j}\right\}
\end{equation*}
See \FigureRef{fig:addp} for an example of adding a point into a mesh pattern.

>>>>

Coincidences among small mesh patterns have been considered by
\textcite{DBLP:journals/corr/ClaessonTU14}, in which the authors use the Simultaneous Shading Lemma,
a closure result and one worked out special case to fully classify coincidences among mesh patterns
of length 2.

@@

We will consider the avoidance sets \(\av{\pi,p}\) where \(\pi\) is a classical pattern of length 3
and \(p\) is a mesh pattern of length 2 in order to establish sufficient conditions for two such
sets to be coincident. We will fix \(\pi\) in order to define these coincidences and say that
\(\pi\)is the \emph{dominating pattern}.

In order to describe the rules it is useful to have a notion for inserting points, ascents, and
descents into a mesh pattern.
\begin{definition}
\label{def:ap}
Let \(p=(\tau,R)\) be a mesh pattern of length \(n\) such that \(\boks{i,j}\notin R\). We define
a mesh pattern \(p^\boks{i,j} = (\tau^\prime,R^\prime)\) of length \(n+1\) as the pattern where a
point is \emph{inserted} into the box \(\boks{i,j}\) in \(G(p)\). Formally the new underlying
classical pattern is defined by
\begin{equation*}
\tau^\prime(k) = \begin{cases}
    j+1 & \text{if } k = i+1\\
    \tau(k) & \text{if } \tau(k)\le j \text{ and }k\le i\\
    \tau(k)+1 & \text{if } \tau(k)> j \text{ and }k\le i\\
    \tau(k-1) & \text{if } \tau(k)\le j \text{ and }k> i+1\\
    \tau(k-1)+1 & \text{if } \tau(k)> j \text{ and }k> i+1\\
\end{cases}
\end{equation*}
While the mesh becomes
\begin{equation*}
\begin{aligned}
R^\prime=&\{\boks{k,\ell}\mid k\le i, \ell\le j, \boks{k,\ell}\in R\} \cup\\
&\{\boks{k,\ell}\mid k\le i, \ell > j, \boks{k,\ell-1}\in R\}\cup\\
&\{\boks{k,\ell}\mid k > i, \ell \le j, \boks{k-1,\ell}\in R\}\cup\\
&\{\boks{k,\ell}\mid k > i, \ell > j, \boks{k-1,\ell-1}\in R\}\\
\end{aligned}
\end{equation*}
\end{definition}
In addition, we give the following definitions:
\begin{definition}
Let \(p=(\tau,R)\) be a mesh pattern of length \(n\) such that \(\boks{i,j}\notin R\) and
\(p^\boks{i,j} = (\tau^\prime,R^\prime)\) to be as defined in \DefinitionRef{def:ap}. We define
the following five modifications of the mesh patterns of the same length as \(p^\boks{i,j}\).
\begin{align*}
p^\boksup{i,j} &= (\tau^\prime, R^\prime \cup \{\boks{i,j+1},\boks{i+1,j+1}\})\\
p^\boksright{i,j} &= (\tau^\prime, R^\prime \cup \{\boks{i+1,j},\boks{i+1,j+1}\})\\
p^\boksdown{i,j} &= (\tau^\prime, R^\prime \cup \{\boks{i,j},\boks{i+1,j}\})\\
p^\boksleft{i,j} &= (\tau^\prime, R^\prime \cup \{\boks{i,j},\boks{i,j+1}\})\\
\end{align*}
\end{definition}
Informally, these are considering the topmost, rightmost, leftmost, or bottommost point in
\(\boks{i,j}\). We allow composition of these modifications, and collect the resulting mesh patterns
in a set
\begin{equation*}
p^\boksall{i,j}=\left\{
p^\boks{i,j},
p^\boksright{i,j},
% p^\boksrightup{i,j},
% p^\boksrightdown{i,j},
p^\boksleft{i,j},
% p^\boksleftup{i,j},
% p^\boksleftdown{i,j},
p^\boksup{i,j},
p^\boksdown{i,j}\right\}
\end{equation*}
See \FigureRef{fig:addp} for an example of adding a point into a mesh pattern.

### 2017-02-20 writing/paper/sections/conclfut.tex
Finish writing paper v1

\chapter{Conclusions and Future work}
\nextvar
% From \ChapterRef{chap:coincs} it can be seen that automatically classifying
% coincidences of mesh patterns is a difficult task, establishing rules for longer
% dominating patterns requires many more cases to be taken.
If we consider a similar system for dominating patterns of length \(4\) and
mesh patterns of length \(2\), it can be seen that the number of cases required
to establish rules increases to a number that is infeasible to compute manually.
For an extension of the First Dominating rule alone, we would have to consider
placement of points in any pair of unshaded regions. The fact that the rules
established do not completely cover the coincidences with a dominating pattern
of length \(3\) (see \TableRef{tab:domclasses}) shows that this is a difficult
task.

It is interesting to consider the application of the Third Dominating rule,
as well as the simple extension of allowing a sequence of \texttt{add\_point}
operations, to mesh patterns without any dominating pattern in order to try to
capture some of the coincidences described in
\textcite{DBLP:journals/combinatorics/HilmarssonJSVU15}
 and \textcite{DBLP:journals/corr/ClaessonTU14}.
 \begin{example}
   We can establish the coincidence between the patterns
   \begin{equation*}
     m_1 = \pattern{}{1,2}{0/1,0/2,1/1,1/2,2/0}, \text{ and } m_2 = \pattern{}{1,2}{0/2,1/0,1/1,2/0,2/1}
   \end{equation*}
   That is not explained by the methods presented by \textcite{DBLP:journals/corr/ClaessonTU14}.

   Consider a permutation containing \(m_1\),
   \begin{equation*}
     \begin{tikzpicture}[scale=0.5]
         \modpattern[5]{}{1,2}{0/1,0/2,1/1,1/2,2/0}
         \draw (1.5,0.5) node {\(Y\)};
         \draw (2.5,1.5) node {\(X\)};
     \end{tikzpicture}
   \end{equation*}
   If the regions corresponding to both \(X\) and \(Y\) are empty then we have
   an occurrence of \(m_2\).
   Consider if the region corresponding to \(X\) is non-empty, we can then choose
   the lowest valued point in this region
   \begin{equation*}
     \begin{tikzpicture}[scale=0.5]
         \modpattern[5]{1,3}{1,3,2}{0/1,0/2,0/3,1/1,1/2,1/3,2/0,2/1,3/0,3/1}
         \draw (1.5,0.5) node {\(Y\)};
     \end{tikzpicture}
   \end{equation*}
   If the region corresponding to \(Y\) is empty then we have an occurrence of
   \(m_2\) with the indicated points.
   Now if the region corresponding to \(Y\) is non-empty, we can choose the
   rightmost point in this region.
   \begin{equation*}
     \begin{tikzpicture}[scale=0.5]
         \modpattern[5]{2,4}{2,1,4,3}{0/2,0/3,0/4,1/2,1/3,1/4,2/0,2/1,2/2,2/3,2/4,3/0,3/1,3/2,4/0,4/1,4/2}
     \end{tikzpicture}
   \end{equation*}
   And now the two indicated points form an occurrence of \(m_2\).
   We have therefore shown that any occurrence of \(m_1\) is an occurrence of
   \(m_2\) and we can easily show the converse by the same reasoning, so \(m_1\)
   and \(m_2\) are coincident.
   This is captured by an extension of the Third Dominating rule where we allow
   multiple steps of adding points before we check for subpattern containment.
 \end{example}
It is not possible to apply the first and second Dominating Rules to the pattern
itself, since when applying the rules we consider containers of the pattern inside
the avoiders of the dominating pattern. For example if we were to attempt to apply
the first rule to the pattern \(\perm{1,2}\) then we would have to consider
containers of \(\perm{1,2}\) inside \(\av{\perm{1,2}}\), and obviously this can
never occur.

It is also possible to take sets of mesh patterns instead of a single mesh pattern
when considering dominating rules, and expressing coincidence between these sets.
Doing this may give nice results. Coincidences between sets with
multiple dominating patterns can also be considered, as this provides even more
power to the rules discussed, these methods may be useful for classifying coincidence
in large sets of pattern. It is also possible to apply these rules to sets where the
dominating pattern is a mesh pattern.

It would be interesting to consider a systematic explanation of Wilf-equivalences
amongst classes where \(\perm{3,2,1}\) is the dominating pattern, possible using the
construction presented in \cite[Sec.~12]{2015arXiv151203226B}, in order to directly
reach enumeration and hopefully establish some of the non-trivial Wilf-equivalences
between classes with different dominating patterns. For example, it is possible to show that
the sets \(\scriptvar = \av{\textpattern{}{1,2}{1/2,2/0,0/0,1/0,1/1,2/1},\perm{2,3,1}}\)\nextvar and
\(\scriptvar = \av{\textpattern{}{1,2}{1/2,2/0,0/0,1/0,1/1,2/1},\perm{3,2,1}}\), are Wilf-equivalent.
This can be seen by considering an occurrence of the mesh pattern,
\(p=\textpattern{}{1,2}{1/2,2/0,0/0,1/0,1/1,2/1}\) in a permutation
\begin{equation*}
    \raisebox{0.6ex}{
    \begin{tikzpicture}[scale=1, baseline=(current bounding box.center)]
        \draw (0.5,1.5) node {\(b_1\)};
        \draw (0.5,2.5) node {\( b_2\)};
        \draw (2.5,2.5) node {\( b_3\)};
        \modpattern[2]{2}{1,2}{0/0,1/0,1/1,1/2,2/0,2/1}
    \end{tikzpicture}
    }
\end{equation*}

If we are inside \(\av{\perm{2,3,1}}\) then any points in the region corresponding
to the box \(b_1\) must be to the right of the points in the region corresponding to
box \(b_2\), and must also form a decreasing subsequence by \LemmaRef{lem:incdec}.
Furthermore, the points in the regions corresponding to \(b_2\) and
\(b_3\) must form an avoider of \(\perm{2,3,1}\) with the indicated point.
Therefore the containers of \(p\) in \(\av{\perm{2,3,1}}\) have generating function
\(\lastvar(x) = x(C(x)-1)/(1-x)\).
Now, if we are in \(\av{\perm{3,2,1}}\) then any points in the region corresponding
to the box \(b_1\) must be to the left of the points in the region corresponding to
box \(b_2\), and must also form a increasing subsequence by \LemmaRef{lem:incdec}.
Furthermore, the points in the regions corresponding to \(b_2\) and
\(b_3\) must form an avoider of \(\perm{3,2,1}\) with the indicated point.
Hence, the containers of \(p\) in \(\av{\perm{3,2,1}}\) have generating function
\(\var(x) = x(C(x)-1)/(1-x)\) and both of these classes have the same
enumeration.

>>>>

If we consider a similar system for dominating patterns of length \(4\) and
mesh patterns of length \(2\), it can be seen that the number of cases required
to establish rules increases to a number that is infeasible to compute manually.
For an extension of the First Dominating rule alone, we would have to consider
placement of points in any pair of unshaded regions. The fact that the rules
established do not completely cover the coincidences with a dominating pattern
of length \(3\) shows that this is a difficult task.

It is interesting to consider the application of the Third Dominating rule, as well as the simple
extension of allowing a sequence of point insertions, to mesh patterns without any dominating
pattern in order to try to capture some of the coincidences described in
\textcite{DBLP:journals/combinatorics/HilmarssonJSVU15} and
\textcite{DBLP:journals/corr/ClaessonTU14}.
 \begin{example}
   We can establish the coincidence between the patterns
   \begin{equation*}
     m_1 = \pattern{}{1,2}{0/1,0/2,1/1,1/2,2/0}, \text{ and } m_2 = \pattern{}{1,2}{0/2,1/0,1/1,2/0,2/1}
   \end{equation*}
   That is not explained by the methods presented by \textcite{DBLP:journals/corr/ClaessonTU14}.

   Consider a permutation containing \(m_1\),
   \begin{equation*}
     \begin{tikzpicture}[scale=0.5]
         \modpattern[5]{}{1,2}{0/1,0/2,1/1,1/2,2/0}
         \draw (1.5,0.5) node {\(Y\)};
         \draw (2.5,1.5) node {\(X\)};
     \end{tikzpicture}
   \end{equation*}
   If the regions corresponding to both \(X\) and \(Y\) are empty then we have
   an occurrence of \(m_2\).
   Consider if the region corresponding to \(X\) is non-empty, we can then choose
   the lowest valued point in this region
   \begin{equation*}
     \begin{tikzpicture}[scale=0.5]
         \modpattern[5]{1,3}{1,3,2}{0/1,0/2,0/3,1/1,1/2,1/3,2/0,2/1,3/0,3/1}
         \draw (1.5,0.5) node {\(Y\)};
     \end{tikzpicture}
   \end{equation*}
   If the region corresponding to \(Y\) is empty then we have an occurrence of
   \(m_2\) with the indicated points.
   Now if the region corresponding to \(Y\) is non-empty, we can choose the
   rightmost point in this region.
   \begin{equation*}
     \begin{tikzpicture}[scale=0.5]
         \modpattern[5]{2,4}{2,1,4,3}{0/2,0/3,0/4,1/2,1/3,1/4,2/0,2/1,2/2,2/3,2/4,3/0,3/1,3/2,4/0,4/1,4/2}
     \end{tikzpicture}
   \end{equation*}
   And now the two indicated points form an occurrence of \(m_2\).
   We have therefore shown that any occurrence of \(m_1\) is an occurrence of
   \(m_2\) and we can easily show the converse by the same reasoning, so \(m_1\)
   and \(m_2\) are coincident.
   This is captured by an extension of the Third Dominating rule where we allow
   multiple steps of adding points before we check for subpattern containment.
 \end{example}

It would be interesting to consider a systematic explanation of Wilf-equivalences among classes
where \(\perm{2,3,1}\) is the dominating pattern, possibly using the  construction presented in
\cite[Sec.~12]{2015arXiv151203226B}, in order to directly reach enumeration and hopefully establish
some of the non-trivial Wilf-equivalences between classes with different dominating patterns.

### 2017-02-20 writing/paper/sections/introduction.tex
Finish writing paper v1

The study of permutation patterns began as a result of Knuth's statements on stack sorting in
\emph{The Art of Computer Programming}\cite[p.~243, Ex.~5,6]{Knuth:1997:ACP:260999}. This original
concept---a subsequence of symbols having a particular relative order, now known as classical
patterns---has been expanded to a variety of definitions. Mesh patterns \textcite{babstein2000} 
considered \emph{vincular} patterns (also known as \emph{generalised} or \emph{dashed} 
patterns) where two adjacent entries in the pattern must also be adjacent in the permutation. 
\textcite{MR2652101} look at classes of pattern where both columns and rows can be shaded, these are
called \emph{bivincular} patterns. \emph{Bruhat-restricted} patterns were studied by 
\textcite{MR2264071} in order to establish necessary conditions for a Schubert variety to be
Gorenstein. All of these definitions are subsumed under the definition of \emph{mesh patterns},
introduced by \textcite{journals/combinatorics/BrandenC11} to capture explicit expansions for 
certain permutation statistics.

When considering permutation patterns some of the main questions posed relate to how and when a
pattern is avoided by, or contained in, a arbitrary set of permutations. Two patterns \(\pi\) and
\(\sigma\) are \emph{Wilf-equivalent} if the number of permutations that avoid \(\pi\) of length 
\(n\) is equal to the number of permutations that avoid \(\sigma\) of length \(n\). A stronger 
equivalence condition is that of \emph{coincidence}, where the set of permutations avoiding \(\pi\)
is exactly equal to the set of permutations avoiding \(\sigma\). Avoiding pairs of patterns of the
same length with certain properties has been studied previously, \textcite{MR2178749} considered 
avoiding a pair of vincular patterns of length 3. \textcite{2015arXiv151203226B} study avoiding a 
vincular and a covincular pattern simultaneously in order to achieve some interesting counting 
results. However, very little work has been done on avoiding a mesh pattern and a classical pattern 
simultaneously.

>>>>

The study of permutation patterns began as a result of Knuth's statements on
stack sorting in \emph{The Art of Computer Programming}\cite[p.~243,
Ex.~5,6]{Knuth:1997:ACP:260999}. This original concept---a subsequence of
symbols having a particular relative order, now known as classical
patterns---has been expanded to a variety of definitions.
\textcite{babstein2000} considered \emph{vincular} patterns (also known as
\emph{generalised} or \emph{dashed} patterns) where two adjacent entries in the
pattern must also be adjacent in the permutation. \textcite{MR2652101} look at
classes of pattern where both columns and rows can be shaded, these are called
\emph{bivincular} patterns. \emph{Bruhat-restricted} patterns were studied by
\textcite{MR2264071} to establish necessary conditions for a Schubert variety to
be Gorenstein. These definitions are subsumed under the definition of \emph{mesh
patterns}, introduced by \textcite{journals/combinatorics/BrandenC11} to capture
explicit expansions for certain permutation statistics.

When considering permutation patterns some of the main questions posed relate to
how and when a pattern is avoided by, or contained in, a arbitrary set of
permutations. Two patterns \(\pi\) and \(\sigma\) are \emph{Wilf-equivalent} if
the number of permutations that avoid \(\pi\) of length \(n\) is equal to the
number of permutations that avoid \(\sigma\) of length \(n\). A stronger
equivalence condition is that of \emph{coincidence}, where the set of
permutations avoiding \(\pi\) is exactly equal to the set of permutations
avoiding \(\sigma\). Avoiding pairs of patterns of the same length with certain
properties has been studied, \textcite{MR2178749} considered avoiding
a pair of vincular patterns of length 3. \textcite{2015arXiv151203226B} study
avoiding a vincular and a covincular pattern simultaneously in order to achieve
some interesting counting results. However, little work has been done on
avoiding a mesh pattern and a classical pattern simultaneously.
